---
title: "Schengen Unemployment"
output:
  word_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r schengen_setup, include=FALSE}
library(tidyverse)
library(readxl)
library(ggpubr)
library(rstatix)
schen_country_list <- c("Austria", "Belgium", "Bulgaria","Croatia","Czechia","Cyprus","Denmark","Estonia","Finland","France","Germany","Greece","Hungary","Iceland","Italy","Latvia","Liechtenstein","Lithuania","Luxembourg","Malta","Netherlands","Norway","Poland","Portugal","Romania","Slovakia","Slovenia","Spain","Sweden","Switzerland")
schen_date_start <-as.Date(c("26-03-1995",	"26-03-1995",	"01-01-2007",	"01-07-2013",	"01-05-2004",	"01-05-2004",	"26-03-1995",	"01-05-2004",	"26-03-1995",	"26-03-1995",	"26-03-1995",	"26-03-1995",	"01-05-2004",	"25-03-2001",	"26-03-1995",	"01-05-2004",	"19-12-2011",	"01-05-2004",	"26-03-1995",	"01-05-2004",	"26-03-1995",	"25-03-2001",	"01-05-2004",	"26-03-1995",	"01-01-2007",	"01-05-2004",	"01-05-2004",	"26-03-1995",	"26-03-1995",	"12-12-2008"), "%d-%m-%Y")
schen_country_date<-tibble(Country = schen_country_list,Date = as.Date(schen_date_start))
my_data <- read_xlsx(path = "~/Schengen_Unemployment/99911-0002_en.xlsx",skip=4,col_names = c("Country","Year","Labour_force",	"Labour_force_participation_rate",	"Rate_of_self-employed_persons",	"Unemployment_rate", "Long-term_unemployment_rate",	"Youth_unemployment_rate",	"Hourly_compensation_costs_in_manufacturing"))%>%
  fill(Country) %>%
  dplyr::select("Country","Year","Labour_force","Labour_force_participation_rate","Unemployment_rate")%>%
  mutate(across(c("Labour_force_participation_rate","Unemployment_rate","Labour_force"),as.double))%>%
  na.omit() %>%
  rowwise() %>%mutate(Zone = 
                        case_when(
                          any(Country==schen_country_date$Country & as.numeric(Year) >= as.numeric(format(schen_country_date$Date,'%Y'))& Country!="Germany") ~ "Schengen",
                          (Country=="Germany") ~ "Domestic",
                          .default = "Non_Schengen"
                        )
  )%>%
  mutate(across(c("Zone","Country"),as.factor))
Mean_rates <- my_data %>%
  group_by(Year,Zone) %>%
  summarise(across(where(is.numeric),mean))
```

Schengen Area Unemployment Comparison

```{r Permit_Show_head, eval=FALSE, echo=TRUE}
head(my_data) # first 6 observation
```
```{r Permit_tell_head,echo=FALSE,message=FALSE}
knitr::kable(head(my_data))# first 6 observation
```
```{r Permit_Show_str, eval=FALSE, echo=TRUE}
str(my_data) # structure of dataset
```
```{r Permit_tell_str,echo=FALSE,message=FALSE}
tibble::glimpse(my_data)
```
```{r Permit_Show_sum, eval=FALSE, echo=TRUE}
summary(my_data) # Summary overview of data
```
```{r Permit_tell_sum,echo=FALSE,message=FALSE}
knitr::kable(summary(my_data)) #Summary overview of data
```
```{r}
UR_plot <- ggplot(Mean_rates, aes(x = Year, y = Unemployment_rate, colour = Zone )) +
  geom_point(size = 2) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90,  hjust=1))+
  facet_wrap(~Zone)
UR_plot

LP_plot <- ggplot(Mean_rates, aes(x = Year, y = Labour_force_participation_rate, colour = Zone )) +
  geom_point(size = 2) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90,  hjust=1))+
  facet_wrap(~Zone)
  LP_plot

```
A visual examination of the data suggests that Mean Unemployment Rate and Labour force participation rate may track more closely to the Domestic (German national) rates. However this could be deceptive as there may more non-schengen states which could have a smoothing effect on the means plotted.

# Analysis

## Unemployment Rate

### Group Summary Statistics

```{r}
my_data %>%
  group_by(Zone)%>%
  get_summary_stats(Unemployment_rate,type = "mean_sd")
ggboxplot(my_data, x="Zone", y= "Unemployment_rate")
```

### Normality Assessment

```{r}
# Build the linear model
model  <- lm(Unemployment_rate ~ Zone, data = my_data)
# Create a QQ plot of residuals
ggqqplot(residuals(model))
```

We can see from the qqplot that the model residuals are not normally distributed, and as such we use the Krushal-Wallis test going forwards, a non-parametric alternative the a one-way ANOVA.


```{r}
res.kruskal <- my_data %>% kruskal_test(Unemployment_rate ~ Zone)
res.kruskal
```

```{r}
 my_data %>% kruskal_effsize(Unemployment_rate ~ Zone)
```



```{r}
# Pairwise comparisons
pwc <- my_data %>% 
  ungroup() %>%
  dunn_test(Unemployment_rate ~ Zone, p.adjust.method = "bonferroni")
pwc
```

```{r}
pwc <- pwc %>% add_xy_position(x = "Zone")
ggboxplot(my_data, x = "Zone", y = "Unemployment_rate") +
  stat_pvalue_manual(pwc, hide.ns = TRUE) +
  labs(
    subtitle = get_test_label(res.kruskal, detailed = TRUE),
    caption = get_pwc_label(pwc)
    )
```
## Labour Force Partcipation Rate

### Group Summary Statistics


```{r}
my_data %>%
  group_by(Zone)%>%
  get_summary_stats(Labour_force_participation_rate,type = "mean_sd")
ggboxplot(my_data, x="Zone", y= "Unemployment_rate")
```

### Normality Assessment

```{r}
# Build the linear model
model  <- lm(Labour_force_participation_rate ~ Zone, data = my_data)
# Create a QQ plot of residuals
ggqqplot(residuals(model))
```

We can see from the qqplot that the model residuals are not normally distributed, and as such we use the Krushal-Wallis test going forwards, a non-parametric alternative the a one-way ANOVA.


```{r}
res.kruskal<- my_data %>% kruskal_test(Labour_force_participation_rate ~ Zone)
res.kruskal
```

```{r}
 my_data %>% kruskal_effsize(Labour_force_participation_rate ~ Zone)
```



```{r}
# Pairwise comparisons
pwc <- my_data %>% 
  ungroup() %>%
  dunn_test(Labour_force_participation_rate ~ Zone, p.adjust.method = "bonferroni")
pwc
```
```{r}
pwc <- pwc %>% add_xy_position(x = "Zone")
ggboxplot(my_data, x = "Zone", y = "Labour_force_participation_rate") +
  stat_pvalue_manual(pwc, hide.ns = TRUE) +
  labs(
    subtitle = get_test_label(res.kruskal, detailed = TRUE),
    caption = get_pwc_label(pwc)
    )
```
